---
title: "HW 10"
author: "Nicholas Alexeev"
date: "11/14/2021"
output: html_document
---
# 1. 1 (a,d) Consider a normal population distribution with the vaue of stdev known
## a. What is confidence level for the interval x +- 2.74 stdev/sqrt(n)
99.4%

## b. What valye of z_(a/2) in the CI formula (7.5) results in a confidence level of 82%?
1.34

# 2. Each of the following is a confidence interval for mean= rue average (i.e., populationmean) resonance frequency (Hz) for all tennis rackets of a certain type:
Type A: (114.4, 115.6)
Type B: (114.1, 115.9)

## a. What is the value of the sample mean resonance frequency for each type?
Type A: (114.4+115.6)/2 = 115.0

Type B: (114.1 + 115.9)/2 = 115.0 

## b. Both intervals were calculated from the same sample data.  The confidence level forone of these intervals is 90% and the other is 99%.  Which of the intervals has the 90%confidence level, and why?  [Hint:  The 99% CI is wider than the 90% CI, but why?  Theanswer doesn’t have anything to do with the formula.  It has to do with what happens99% of the time versus 90% of the time.]
Type A has the 99% confidence interval because it has a wider spread

# 3. modified.   Assume  that  the  helium  porosity  (in  percentage)  of  coal  samplestaken  from  any  particular  seam  is  normally  distributed  with  true  standard  deviation 0.75.
## a. Compute  a  95%  CI  for  the  true  average  porosity  of  a  certain  seam  if  the  average porosity for 20 specimens from the seam was 4.85.
$$
\left(4.85-1.96*\frac{0.75}{\sqrt{20}},4.85+1.96*\frac{0.75}{\sqrt{20}}\right) = \left(4.85-0.33,4.85+0.33\right) = (4.52,5.18)
$$

## b. Provide an informal interpretation of the CI in (a).
There is a 95% chance that the true mean lies inside of the interval (4.52,5.18).

## c. Provide a formal interpretation of the CI in (a).
If the experiment is repeated, there is a 95% chance that the measured sample mean will line in the interval (4.52,5.18).

## d. What sample size is necessary to estimate true average porosity to within 0.2 with 99% confidence?
$$
0.2 = 2.58*\frac{0.75}{\sqrt{n}}
$$
$$
\sqrt{n} = \frac{0.75*2.58}{0.2} = 96
$$

$$
n = 94
$$

# 4. 7.2  #  18  (modified)  The  U.S.  Army  commissioned  a  study  to  assess  how  deeply  abullet penetrates ceramic body armor.  (Reference cited in text.)  In the standard test,a  cylindrical  clay  model  is  layered  under  the  armor  vest.   A  projectile  is  then  fired,causing an indentation in the clay.  The deepest impression in the clay is measured asan indication of survivability of someone wearing the armor.  The data for the problemappears in the text; I’ve also posted a file with the data and also R code for this problem.The measurements are in mm.

## a. Construct a boxplot and a histogram of the data and comment on features.  Be sureto include the plots in your solutions
```{r}
Depth <- c(
22.4,23.6,24.0,24.9,25.5,25.6,
25.8,26.1,26.4,26.7,27.4,27.6,
28.3,29.0,29.1,29.6,29.7,29.8,
29.9,30.0,30.4,30.5,30.7,30.7,
31.0,31.0,31.4,31.6,31.7,31.9,
31.9,32.0,32.1,32.4,32.5,32.5,
32.6,32.9,33.1,33.3,33.5,33.5,
33.5,33.5,33.6,33.6,33.8,33.9,
34.1,34.2,34.6,34.6,35.0,35.2,
35.2,35.4,35.4,35.4,35.5,35.7,
35.8,36.0,36.0,36.0,36.1,36.1,
36.2,36.4,36.6,37.0,37.4,37.5,
37.5,38.0,38.7,38.8,39.8,41.0,
42.0,42.1,44.6,48.3,55.0)
boxplot(Depth)
hist(Depth)
```
The plots resemble a normal distribution except the depth is skewed to the right.

## b. Construct  a  normal  probability  plot,  and  include  it  with  your  solutions.   (Again,the  R  code  is  in  the  attached  file.)   Is  it  plausible  that  impression  depth  is  normally distributed?  Is a normal distribution assumption needed in order to calculate a CI orbound for the true average depth using the included data?  Explain
```{r}
library(car)
qqPlot(Depth)
```
It is not reasonable to assume that the data is normally distributed as the high values do not follow a normnal distribution.

## c. Theinterpretation of this type of interval is this:  We are 99% confident that the true average impression depth lies below ____.
```{r}
length(Depth) 
qt(0.995,83) 
```

```{r}
mean(Depth)-qt(0.995,83) *sd(Depth)/sqrt(length(Depth))
```
```{r}
t.test(Depth,conf.level=0.99)
```
We are 99% confident that the true average impression depth lies below 32.02.

# 5. In a Quinnipiac (KWIN-uh-pea-ack) University poll in late September, 2021, a sample ofapproximately 1000 registered Texas voters were asked whether abortion should be legalin cases of rape or incest.  Of these, approximately 77% said Yes; 16% said No, and 7%had other responses.  Using these figures as if they are exact (770 out of 1000 said Yes),construct a 95% confidence interval for the true proportion of all Texas registered voterswho would answer Yes to this question.  Interpret your CI. (Use the better of the twoformulas discussed in class, on page 168 of the lecture notes.)
$$
\sigma = \sqrt{\frac{p(1-p)}{n}} = \sqrt{\frac{0.77*0.23}{1000}} = 0.013
$$
$$
(0.77-1.96*0.013,0.77+1.96*0.013) = (0.744,0.796)
$$

# 6.  Determine the values of the following quantities; use both R (the function qt) and TableA.5.

## a. t_0.1,df=15
1.341
```{r}
qt(1-0.1,15)
```

## b. t_0.05 df = 31
1.695
```{r}
qt(1-0.05,31)
```

## c. t_0.025 df = 40.
2.021
```{r}
qt(1-0.025,31)
```


# 7. Find  the t critical  value  for  a  two-sided  confidence  interval  in  each  of  the  following situations; use R and also Table A.5.
## a. df = 10, 95% COnfidence level.
2.228
```{r}
qt(0.95+0.025,10)
```

## b. df = 15. 90% confidence level.
1.753
```{r}
qt(0.95,15)
```

## c. df = 45, 99% confidence level.
2.691
```{r}
qt(0.995,45)
```

# 8. Use the cadence data from problem # 37 in 7.3 to form a 95% confidence interval for theaverage cadence using the bootstrap method.  Compare this with the interval we foundin class (page 172 of the lecture notes).  R code for the bootstrap procedure appears onpage 165 of the lecture notes.
```{r}
data <- c(0.95,0.85,0.92,0.95,0.93,0.86,1.0,0.92,0.85,0.81,0.78,0.93,0.93,1.05,0.93,1.06,0.96,0.81,0.96)
n <- length(data)
b <- 5000
results <- rep(NA,b)
for(i in 1:b){
  my.sample <- sample(data,size = n, replace = TRUE)
  results[i]<-mean(my.sample)
}
length(unique(results))
lower.lim <- quantile(results,0.025)
upper.lim <- quantile(results,0.975)
c(lower.lim,upper.lim)
```
I dont 
# 9. In an  FBI study on  propery crimes,  a  sample of 25  pickpocket offenses was  analyzed.The data (amount stolen) appear to come from a normal population.  The mean amountstolen was$513.32 and the sample standard deviation was$262.23.  Find and interpret a 95% confidence interval for the average amount stolen by pickpockets
$$
\left(513.32-2.064*\frac{262.23}{\sqrt{25}},513.32+2.064*\frac{262.23}{\sqrt{25}}\right) = (513.32-108,513.32+108) = (404.98,621.47)
$$

# 10. For  the  following  pairs  of  hypotheses,  indicate  whether  each  does  or  does  not  comply with our rules for setting up hypotheses.  If a pair doesn’t comply with the rules, be sure to say why not.
## a. H0, u = 50, Ha u <50
Yes

## b. H0 p != 0.6 vs Ha p=0.6
invalid, H0 must be an equality

## c. H0 p_hat = 0.5 vs Ha p_hat != 0.5
No as p_hat describes a sample rather then the population

## d. H0 u = 5.1 vs Ha u > 5.4
invalid there is a gap between H0 and Ha.

## e. H0 x_bar = 16.0 vs Ha x_bar != 16.0
No as x_bar describes a sample rather then the population

## f. H0 u1 -u2 = 10.0 vs Ha u1-u2 !=10.0
yes
